---
title: "Analysis"
author: "ZhijianLiu"
date: "4/19/2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
#1.H1B for DC, VA, MD
```{r read}
h1b <- readRDS("h1b.rds") 
```
##1.1 Annual H1B line plot (2011 - 2016)
```{sub 3 states}
sub_h1b <- readRDS("h1b.rds") %>%
  filter(state == "DISTRICT OF COLUMBIA" |state == "MARYLAND" | state == "VIRGINIA")
```
###1.1.1 application

###1.1.2 salary

###1.1.3 how they rank in the whole country(2011- 2016)

###1.2 city ranks in these three states

###1.3 employer ranks in these three states

##2 data science related
###2.1 number of application of different data science job change through years(line plot)

###2.2 rank of number of h1b application of data science related jobs in all states( compare with h1b means)

###2.3 salary of data science related jobs in all states( compare with h1b means)

###2.4 salary of different data science job

###2.5 certificaion rate





###salary
h1b %>% 
+ filter (state == "SOUTH DAKOTA") %>%
+ arrange(desc(PREVAILING_WAGE))
