---
title: "tidy"
author: "Zhijian Liu"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(magrittr)
library(tidyverse)
```

#1.Downlaod and output
```
if(!file.exists("./data")) {dir.create("./data")}
download.file("https://www.dropbox.com/s/qalyqxdsnx39343/h1b_tidy.csv?dl=1", destfile = "./h1b1.csv", mode = "wb")
h1b <- read_csv("./h1b1.csv")
h1b <- h1b %>%
  select(-1)
write.csv(h1b, "./data/h1b_tidy.csv")
saveRDS(h1b,"h1b.rds")
file.remove("./h1b1.csv")
```

---
#2.Tidy
```{r read}
h1b <- readRDS("./h1b.rds") %>%
  select(CASE_STATUS:WORKSITE, lon, lat)
```
##2.1wrong worksite
###2.1.1wrong format, start with "#"
```{r "#"}
h1b1 <- h1b
#detect
h1b1 %>%
  filter(str_detect(h1b1$WORKSITE, "#")) %>%
  select(WORKSITE) %>%
  unlist() %>%
  unname() %>%
  head()
#replace test
h1b1$WORKSITE <- str_replace(h1b1$WORKSITE, ".*#.*,","unknown,")
#test
sum(str_detect(h1b1$WORKSITE, "#"))
sum(str_detect(h1b$WORKSITE, "#"))
#replace 
h1b$WORKSITE <- str_replace(h1b$WORKSITE, ".*#.*,","unknown,")
```
###2.1.2with digit
```{r digit}
#detect
h1b %>%
  filter(str_detect(h1b1$WORKSITE, "\\d")) %>%
  select(WORKSITE) %>%
  unlist() %>%
  unname() %>%
  head()
#replace test
h1b1$WORKSITE <- str_replace(h1b1$WORKSITE, ".*\\d.*,","unknown,")
#test
sum(str_detect(h1b1$WORKSITE, "\\d"))
sum(str_detect(h1b$WORKSITE, "\\d"))
#replace 
h1b$WORKSITE <- str_replace(h1b$WORKSITE, ".*\\d.*,","unknown,")
```
###2.1.3only state, no city, start with ","
```{r ","}
#detect
h1b %>%
  filter(str_detect(h1b1$WORKSITE, "^,")) %>%
  select(WORKSITE) %>%
  unlist() %>%
  unname()
#replace test
h1b1$WORKSITE <- str_replace(h1b1$WORKSITE, "^,","unknown,")
#test
sum(str_detect(h1b1$WORKSITE, "^,"))
sum(str_detect(h1b$WORKSITE, "^,"))
#replace 
h1b$WORKSITE <- str_replace(h1b$WORKSITE, "^,","unknown,")
```
```{r save}
saveRDS(h1b,"h1b.rds")
rm(list=ls())
```

##2.2 seperate and transfer
```{r read}
h1b <- readRDS("./h1b.rds")
```
###2.2.1 seperate and transfer unknown city
```{r seperate, transfer unknown city to NA}
#seperate "WORKSITE"" into "city"" and "state"
h1b <- h1b %>%
  separate(WORKSITE,c("city","state"),", ",remove = FALSE)
#detect "unknow"
h1b %>%
  filter(city == "unknown") %>%
  head()
#replace "unkown" with NA
h1b$city[h1b$city == "unknown"] <- NA 
#test
sum(is.na(h1b$city))
h1b %>%
  filter(city == "unknown") %>%
  nrow()
```
###2.2.2 transfer unknown WORKSITE
```{r transfer unknown WORKSITE to NA}
#detect "unknow"
h1b %>%
  filter(str_detect(h1b$WORKSITE, ".*unknown.*,.*")) %>%
  nrow()
#replace "unkown" with NA
h1b$WORKSITE <- gsub(".*unknown.*,.*", NA, h1b$WORKSITE)
#test
sum(is.na(h1b$WORKSITE))
h1b %>%
  filter(str_detect(h1b$WORKSITE, ".*unknown.*,.*")) %>%
  nrow()
```
```{r newyork}
h1b$WORKSITE <- str_replace(h1b$WORKSITE, "NEWYORK, NEW YORK","NEW YORK, NEW YORK")
```
```{r save}
saveRDS(h1b,"h1b.rds")
rm(list=ls())
```

---
#3.new variables for analysis
```{r read}
h1b <- readRDS("./h1b.rds")
```
##3.1 H1B application
###3.1.1 state application rank
```{r state_application}
count_state <- h1b %>%
  group_by(state) %>%
  summarise(apply_state = sum(!is.na(CASE_STATUS))) %>%
  arrange(desc(apply_state)) %>%
  mutate( applyrank_state = min_rank (desc(apply_state)))
count_state

h1b <- h1b %>%
  full_join(count_state) #join h1b dataset
```
###3.1.2 city application rank
```{r city_application}
#country_wide
count_site <- h1b %>%
  group_by(WORKSITE) %>%
  summarise(apply_city = sum(!is.na(CASE_STATUS)|!is.na(WORKSITE))) %>%
  arrange(desc(apply_city)) %>%
  mutate( applyrank_city_us = min_rank (desc(apply_city)))
count_site
saveRDS(count_site,"count_site.rds") #for ggmap

h1b <- h1b %>%
  left_join(count_site, by = "WORKSITE") #join h1b dataset

#within state
count_city_state <- h1b %>%
  group_by(state, city) %>%
  summarise(apply_city = sum(!is.na(CASE_STATUS))) %>%
  group_by(state) %>%
  mutate( applyrank_city_state = min_rank (desc(apply_city)))

count_city_state <- count_state %>% 
  full_join(count_city_state, by = "state") %>%
  arrange(applyrank_state, applyrank_city_state) %>%
  select(state, city, applyrank_city_state)
count_city_state

h1b <- h1b %>%
  left_join(count_city_state, by = c("city","state"))#join h1b dataset
```
###3.1.3 rank of employers
```{r employer_application}
#country_wide
count_employer_us <- h1b %>%
  group_by(EMPLOYER_NAME) %>%
  summarise(apply_employer_us = sum(!is.na(CASE_STATUS)|!is.na(EMPLOYER_NAME))) %>%
  arrange(desc(apply_employer_us)) %>%
  mutate( applyrank_employer_us = min_rank (desc(apply_employer_us)))
count_employer_us

h1b <- h1b %>%
  left_join(count_employer_us, by = "EMPLOYER_NAME")#join h1b dataset

#within state
count_employer_state <- h1b %>%
  group_by(state, EMPLOYER_NAME) %>%
  summarise(apply_employer_state = sum(!is.na(CASE_STATUS)|!is.na(EMPLOYER_NAME))) %>%
  group_by(state) %>%
  mutate(applyrank_employer_state = min_rank (desc(apply_employer_state)))

count_employer_state <- count_state %>% 
  full_join(count_employer_state, by = "state") %>%
  arrange(applyrank_state, applyrank_employer_state) 
count_employer_state

h1b <- h1b %>%
  left_join(count_employer_state[c("state","EMPLOYER_NAME","applyrank_employer_state","apply_employer_state")],
            by = c("state","EMPLOYER_NAME"))#join h1b dataset

#within city
count_employer_city <- h1b %>%
  group_by(WORKSITE, EMPLOYER_NAME) %>%
  summarise(apply_employer_city = sum(!is.na(CASE_STATUS)|!is.na(EMPLOYER_NAME)|!is.na(WORKSITE))) %>%
  group_by(WORKSITE) %>%
  mutate( applyrank_employer_city = min_rank (desc(apply_employer_city)))

count_employer_city <- count_site %>% 
  full_join(count_employer_city, by = "WORKSITE") %>%
  arrange(applyrank_city_us, applyrank_employer_city) 
count_employer_city

h1b <- h1b %>%
  left_join(count_employer_city[c("WORKSITE","EMPLOYER_NAME","apply_employer_city","applyrank_employer_city")],
            by = c("WORKSITE","EMPLOYER_NAME"))#join h1b dataset
```
###3.1.4 rank of job
```
There are too many occupational names, it would take a lot of effort to classify all of them. Here i will only analyze data science ralated jobs.
```

##3.2 H1B salary
```{r overview}
h1b %>%
  arrange(desc(PREVAILING_WAGE)) %>%
  filter(CASE_STATUS == "CERTIFIED") %>%
  head()
h1b %>%
  arrange(desc(PREVAILING_WAGE)) %>%
  head()
#reasonable subset for analysis
h1b_salary <- h1b %>%
  filter(PREVAILING_WAGE <= 400000000|is.na(PREVAILING_WAGE))
```
###3.2.1 state salary rank
```{r state_salary}
salary_state <- h1b_salary %>%
  group_by(state) %>%
  summarise(wage_state = mean(PREVAILING_WAGE, na.rm = TRUE, trim = 0.005)) %>% #get rid of extrem value
  arrange(desc(wage_state)) %>%
  mutate( wagerank_state = min_rank (desc(wage_state)))
salary_state

h1b <- h1b %>%
  left_join(salary_state, by = "state")#join h1b dataset

```
###3.2.2 city salary rank
```{r city_salary}
#country_wide
salary_site <- h1b_salary %>%
  group_by(WORKSITE) %>%
  summarise(wage_city = mean(PREVAILING_WAGE, na.rm = TRUE, trim = 0.05),
            apply_city_us = sum(!is.na(CASE_STATUS)|!is.na(EMPLOYER_NAME))) %>%
  filter(apply_city_us >= 50) %>%
  arrange(desc(wage_city)) %>%
  mutate( wagerank_city_us = min_rank (desc(wage_city)))
salary_site

h1b <- h1b %>%
  left_join(salary_site[c("wage_city","WORKSITE","wagerank_city_us")], by = "WORKSITE")#join h1b dataset

#within state
salary_city_state <- h1b_salary %>%
  group_by(state, city) %>%
  summarise(wage_city_state = mean(PREVAILING_WAGE, na.rm = TRUE, trim = 0.05),
            apply_city_us = sum(!is.na(CASE_STATUS)|!is.na(EMPLOYER_NAME))) %>%
  filter(apply_city_us >= 50) %>%
  group_by(state) %>%
  mutate( wagerank_city_state = min_rank (desc(wage_city_state)))

salary_city_state <- salary_state %>% 
  full_join(salary_city_state, by = "state") %>%
  arrange(wagerank_state, wagerank_city_state) 
salary_city_state

h1b <- h1b %>%
  left_join(salary_city_state[c("city","state","wagerank_city_state","wage_city_state")], 
            by = c("city","state"))#join h1b dataset

```
###3.2.3 employer salary rank
```{r employer salary}
#country_wide
salary_employer_us <- h1b_salary %>%
  group_by(EMPLOYER_NAME) %>%
  summarise(wage_employer_us = mean(PREVAILING_WAGE, na.rm = TRUE),
            apply_employer_us = sum(!is.na(CASE_STATUS)|!is.na(EMPLOYER_NAME))) %>%
  filter(apply_employer_us >= 30) %>%
  arrange(desc(wage_employer_us)) %>%
  mutate( wagerank_employer_us = min_rank (desc(wage_employer_us)))
salary_employer_us

h1b <- h1b %>%
  left_join(salary_employer_us[c("wage_employer_us","EMPLOYER_NAME","wagerank_employer_us")], 
            by = "EMPLOYER_NAME")#join h1b dataset

#within state
salary_employer_state <- h1b_salary %>%
  group_by(state, EMPLOYER_NAME) %>%
  summarise(wage_employer_state = mean(PREVAILING_WAGE, na.rm = TRUE),
            apply_employer_state = sum(!is.na(CASE_STATUS)|!is.na(EMPLOYER_NAME))) %>%
  filter(apply_employer_state >= 50) %>%
  group_by(state) %>%
  mutate(wagerank_employer_state = min_rank (desc(wage_employer_state)))

salary_employer_state <- salary_state %>% 
  full_join(salary_employer_state, by = "state") %>%
  arrange(wagerank_state, wagerank_employer_state) 
salary_employer_state

h1b <- h1b %>%
  left_join(salary_employer_state[c("wage_employer_state","EMPLOYER_NAME","wagerank_employer_state","state")], 
            by = c("EMPLOYER_NAME","state")) #join h1b dataset

#within city
salary_employer_city <- h1b_salary %>%
  group_by(WORKSITE, EMPLOYER_NAME) %>%
  summarise(wage_employer_city = mean(PREVAILING_WAGE, na.rm = TRUE),
            apply_employer_city = sum(!is.na(CASE_STATUS)|!is.na(EMPLOYER_NAME))) %>%
  filter(apply_employer_city >= 10) %>% 
  group_by(WORKSITE) %>%
  mutate( wagerank_employer_city = min_rank (desc(wage_employer_city)))
salary_employer_city

h1b <- h1b %>%
  left_join(salary_employer_city[c("wage_employer_city","EMPLOYER_NAME","wagerank_employer_city","WORKSITE")], 
            by = c("EMPLOYER_NAME","WORKSITE"))#join h1b dataset
```
```{r save}
names(h1b)
saveRDS(h1b,"h1b.rds")
rm(list=ls())
```
